Causal Inference with Graphical Neural Networks

### 2024-04-12 latex/draft.tex
Pushing latex changes.

\section{Background}

@@

>>>>

Each node of the graph was is associated with a specific conditional probability that the
creators of the dataset used to generate the data. These probabilities can be found in Table~\ref{tab:lucas_probs}
in Appendix~\ref{create}.

\section{do Calculus Result stuff (idk where this goes)}

Let's consider the following conditional probabilities with \textbf{do} operators applied:

\begin{align*}
    P(\text{LC} = \text{T} \;&|\; \textbf{do}(\text{YF} = \text{T})), \\
    P(\text{LC} = \text{T} \;&|\; \textbf{do}(\text{PP} = \text{T})), \\
    P(\text{LC} = \text{T} \;&|\; \textbf{do}(\text{A} = \text{T})), \\
    P(\text{LC} = \text{T} \;&|\; \textbf{do}(\text{AD} = \text{T})), \\
    P(\text{LC} = \text{T} \;&|\; \textbf{do}(\text{CA} = \text{T})), \\
\end{align*}

where LC is lung cancer, YF is yellow fingers, PP is peer pressure, A is anxiety, AD is attention disorder, and CA 
is car accident. We note that we chose this probability because it does not involve any
variables that are in the Markov blanket of our target variable. Thus, we should expect that there is very little
predictive power in this probability.

Plugging forward with \textbf{do} calculus, we can write $P(\text{LC} = \text{T} \;|\; \textbf{do}(\text{YF} = \text{T}))$ as
(using corollary~\ref{corollary:do})

\begin{equation*}
    P(\text{LC} = \text{T} \;|\; \textbf{do}(\text{YF} = \text{T})) = 
    \sum_{s\in\{T, F\}} \frac{P(\text{YF} = \text{T},\, \text{LC} = \text{T},\, \text{S} = s)}{P(\text{YF} = \text{T}\,|\, \text{S} = s)},
\end{equation*}

where $\text{S}$ is smoking. Using the property that we can split the numerator into probabilities only involving parents ({\color{red}rewrite})

\begin{equation*}
    \sum_{s\in\{T, F\}}\frac{P(\text{YF} =\text{T}\;|\; \text{S} = s) P(\text{LC} 
    = \text{T}\;|\; \text{S}=s,\, \text{G}) P(\text{S}=s\;|\;\text{A},\, \text{PP})}
    {P(\text{YF} = \text{T}\;|\; \text{S} = s)}.
\end{equation*}

All of these fun conditionals lead us to

\begin{equation*}
    \sum_{s,a,g,p}\frac{P(\text{YF} =\text{T}\;|\; \text{S} = s) P(\text{LC} 
    = \text{T}\;|\; \text{S}=s,\, \text{G}=g) P(\text{S}=s\;|\;\text{A}=a,\, \text{PP}=p)}
    {P(\text{YF} = \text{T}\;|\; \text{S} = s)},
\end{equation*}

where each $s,a,g,p$ is summed over the set $\{T, F\}$.

@@

\newpage

\section*{Appendix}

\appendix
\section{Creating the LUCAS0 Dataset}
\label{create}

As described in the \nameref{data} section, the LUCAS0 dataset is a toy dataset generated 
artificially by causal Bayesian networks with binary variables. The dataset consists of 
11 nodes and 2000 training samples. The causal diagram for the LUCAS0 dataset is shown in Figure~\ref{fig:lucas_dag}.

On the website, the authors mention that each node of the graph is associated with conditional probabilities
which were used to generate the data. These probabilities are found in Table~\ref{tab:lucas_probs}.

\begin{table}[h]
    \centering
    \begin{tabular}{|c|c|}
        \hline
        \multicolumn{2}{|c|}{\textbf{Conditional Probabilities}} \\
        \hline
        \textit{Conditional} & \textit{Probability} \\
        \hline
        P(Anxiety=T) & 0.64277 \\ \hline
        P(Peer Pressure=T) & 0.32997 \\ \hline
        P(Smoking=T $\vert$ Peer Pressure=F, Anxiety=F) & 0.43118 \\ \hline
        P(Smoking=T $\vert$ Peer Pressure=T, Anxiety=F) & 0.74591 \\ \hline
        P(Smoking=T $\vert$ Peer Pressure=F, Anxiety=T) & 0.8686 \\ \hline
        P(Smoking=T $\vert$ Peer Pressure=T, Anxiety=T) & 0.91576 \\ \hline
        P(Yellow Fingers=T $\vert$ Smoking=F) & 0.23119 \\ \hline
        P(Yellow Fingers=T $\vert$ Smoking=T) & 0.95372 \\ \hline
        P(Genetics=T) & 0.15953 \\ \hline
        P(Lung cancer=T $\vert$ Genetics=F, Smoking=F) & 0.23146 \\ \hline
        P(Lung cancer=T $\vert$ Genetics=T, Smoking=F) & 0.86996 \\ \hline
        P(Lung cancer=T $\vert$ Genetics=F, Smoking=T) & 0.83934 \\ \hline
        P(Lung cancer=T $\vert$ Genetics=T, Smoking=T) & 0.99351 \\ \hline
        P(Attention Disorder=T $\vert$ Genetics=F) & 0.28956 \\ \hline
        P(Attention Disorder=T $\vert$ Genetics=T) & 0.68706 \\ \hline
        P(Born an Even Day=T) & 0.5 \\ \hline
        P(Allergy=T) & 0.32841 \\ \hline
        P(Coughing=T $\vert$ Allergy=F, Lung cancer=F) & 0.1347 \\ \hline
        P(Coughing=T $\vert$ Allergy=T, Lung cancer=F) & 0.64592 \\ \hline
        P(Coughing=T $\vert$ Allergy=F, Lung cancer=T) & 0.7664 \\ \hline
        P(Coughing=T $\vert$ Allergy=T, Lung cancer=T) & 0.99947 \\ \hline
        P(Fatigue=T $\vert$ Lung cancer=F, Coughing=F) & 0.35212 \\ \hline
        P(Fatigue=T $\vert$ Lung cancer=T, Coughing=F) & 0.56514 \\ \hline
        P(Fatigue=T $\vert$ Lung cancer=F, Coughing=T) & 0.80016 \\ \hline
        P(Fatigue=T $\vert$ Lung cancer=T, Coughing=T) & 0.89589 \\ \hline
        P(Car Accident=T $\vert$ Attention Disorder=F, Fatigue=F) & 0.2274 \\ \hline
        P(Car Accident=T $\vert$ Attention Disorder=T, Fatigue=F) & 0.779 \\ \hline
        P(Car Accident=T $\vert$ Attention Disorder=F, Fatigue=T) & 0.78861 \\ \hline
        P(Car Accident=T $\vert$ Attention Disorder=T, Fatigue=T) & 0.97169 \\ \hline
        \hline
    \end{tabular}
    \caption{Conditional probabilities for the nodes in the LUCAS0 dataset.}
    \label{tab:lucas_probs}
\end{table}
